---
title: "CP1 - ToothGrowth"
author: "Petar Luketic"
date: "Sunday, June 21, 2015"
output:
  pdf_document: default
  html_document:
    theme: journal
---

# Tooth Growth 
  
## 1. Load the ToothGrowth data and perform some basic exploratory data analysis     
  
```{r explore, echo=TRUE}
# load the dataset
library(datasets)
data(ToothGrowth)

# look at the dataset variables
head(ToothGrowth)
str(ToothGrowth)

# convert variable dose from numeric to factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# review dataset variables after conversion
str(ToothGrowth)

# number of rows of dataset
nrow(ToothGrowth)
```  
  
## 2. Provide a high level summary of the data.  
  
```{r Summary, echo=TRUE}  
# summary statistics for all variables
summary(ToothGrowth)

# split cases on different dose levels and delivery methods
table(ToothGrowth$dose, ToothGrowth$supp)

library(ggplot2)
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity",) +
    facet_grid(. ~ supp) +
    xlab("Dose in miligrams") +
    ylab("Tooth length") +
    guides(fill=guide_legend(title="Supplement type"))
```  
  
For both delivery methods, there is positive correlation between the tooth length and the dose levels of Vitamin C.    
  
## 3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.  
  
95% confidence intervals for two variables and the intercept are as follows:  

```{r ci, echo=TRUE}
fit <- lm(len ~ dose + supp, data=ToothGrowth)
confint(fit)
```  
  
The confidence intervals suggest that 95% of the time, the coefficient estimations will be in these ranges. For each coefficient (i.e. intercept, `dose` and `suppVC`), the null hypothesis should be that the coefficients are zero which means that no tooth length variation is explained by that variable. 
  
 
```{r}
summary(fit)
```  
  
The model explains 70% of the data variance. The intercept is `r fit$coefficients[[1]]`, so without Vitamin C, the average tooth length is `r fit$coefficients[[1]]`. The coefficient of `dose` is `r fit$coefficients[[2]]`. It can be interpreted as increasing the delievered dose 1 mg, all else equal increases the tooth length `r fit$coefficients[[2]]`. 
